pdp: An R Package for Constructing Partial Dependence Plots

### 2016-12-28 manuscript/greenwell.tex
fixed bug in pdRegression.xgb.Booster

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=1.0\linewidth]{partial_boston_ice}
  \caption{ICE plots and PDP for the Boston housing data. \textit{Left}: ICE curves. \textit{Middle}: Specific quantiles of ICE curves. \textit{Right}: Standard PDP.}
  \label{fig:partial_boston_ice}
\end{figure}

@@

# Grid of tuning values for XGBoost.
xgb.grid <- expand.grid(nrounds = c(100, 500, 1000, 2000, 5000),
                        max_depth = 1:6,
                        eta = c(0.001, 0.01, 0.1, 0.5, 1),
                        gamma = 0,
                        colsample_bytree = 1,
                        min_child_weight = 1)

# Train an XGBoost model
set.seed(202)
boston.xgb <- train(x = data.matrix(subset(boston, select = -cmedv)),
                    y = boston$cmedv,
                    method = "xgbTree",
                    metric = "RMSE",
                    trControl = ctrl,
                    tuneGrid = xgb.grid)
\end{example}
The model with the smallest root-mean-squared error (RMSE) had \code{nrounds = 2000}, \code{max\_depth = 3}, \code{eta = 0.01}, and achieved a cross-validated RMSE and $R^2$ of 2.819 and 0.905, respectively. The following snippet of code computes the partial dependence of \code{cmedv} on both \code{rm} and \code{lstat}, individually and together, and then uses \code{grid.arrange} to display all three PDPs on the same page. The results are displayed in Figure~\ref{fig:boston_xgb}.
\begin{example}
grid.arrange(
  # Figure 8 (left)
  partial(boston.xgb, pred.var = "rm", plot = TRUE, rug = TRUE),
  # Figure 8 (middle)
  partial(boston.xgb, pred.var = "lstat", plot = TRUE, rug = TRUE),
  # Figure 8 (right)
  partial(boston.xgb, pred.var = c("lstat", "rm"), plot = TRUE, chull = TRUE),
  ncol = 3
)
\end{example}

>>>>

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=1.0\linewidth]{partial_boston_ice}
  \caption{ICE curves and PDP depicting the relationship between \code{cmedv} and \code{lstat} for the Boston housing example. \textit{Left}: ICE curves. \textit{Middle}: Specific quantiles of ICE curves. \textit{Right}: Standard PDP.}
  \label{fig:partial_boston_ice}
\end{figure}

@@

# Grid of tuning values for XGBoost.
xgb.grid <- expand.grid(nrounds = c(1000, 2000, 3000),
                        max_depth = 1:4,
                        eta = c(0.001, 0.01, 0.1),
                        gamma = 0,
                        colsample_bytree = 1,
                        min_child_weight = 1,
                        subsample = c(0.5, 0.75, 1))

# Train an XGBoost model
set.seed(202)
boston.xgb <- train(x = data.matrix(subset(boston, select = -cmedv)),
                    y = boston$cmedv,
                    method = "xgbTree",
                    metric = "RMSE",
                    trControl = ctrl,
                    tuneGrid = xgb.grid)
\end{example}
The model with the smallest root-mean-squared error (RMSE) had \code{nrounds = 3000}, \code{max\_depth = 3}, \code{eta = 0.01}, and \code{subsample = 0.75}; the final model achieved a cross-validated RMSE and $R^2$ of 2.789 and 0.906, respectively. The following snippet of code computes the partial dependence of \code{cmedv} on both \code{rm} and \code{lstat}, individually and together, and then uses \code{grid.arrange} to display all three PDPs on the same page. The results are displayed in Figure~\ref{fig:boston_xgb}.
\begin{example}
grid.arrange(
  # Figure 8 (left)
  partial(boston.xgb, pred.var = "rm", plot = TRUE, rug = TRUE),
  # Figure 8 (middle)
  partial(boston.xgb, pred.var = "lstat", plot = TRUE, rug = TRUE),
  # Figure 8 (right)
  partial(boston.xgb, pred.var = c("lstat", "rm"), plot = TRUE, chull = TRUE),
  ncol = 3
)
\end{example}
